---
title: "Base R Programming"
format: html
editor: visual
---

# Task 1: Basic Vector Practice

1.  Creating pre- and post-treatment Vectors:

```{r}
pre <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

post <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

2.  Now generate and assign names to each:

```{r}
names <- paste("Subject", 1:20, sep = "_")

names(pre) <- names
names(post) <- names

names(pre)
```

3.  Find **decrease** in blood pressure:

```{r}
dec <- pre-post

dec
```

4.  Find the average **decrease** in blood pressure:

```{r}
ave_dec <- mean(dec)

ave_dec
```

5.  Determine which patients experienced a decrease:

```{r}
which(dec > 0)
```

6.  Subset the vector of differences with only positive changes:

```{r}
pos_dec <- dec[dec>0]

pos_dec
```

7.  Calculate average positive change:

```{r}
ave_pos_dec <- mean(pos_dec)

ave_pos_dec
```

# Task 2: Basic Data Frame Practice

1.  Create a df:

```{r}
df <- data.frame(patient=names(pre), pre_bp=pre, post_bp=post, diff_bp=dec, row.names=NULL)

df
```

2.  Get only rows of *df* where change is negative:

```{r}
df[df$diff_bp<0,]
```

3.  Add new boolean column (*post_bp* \< 120):

```{r}
df$low_post_bp <- df$post_bp<120

df
```

4.  Print df:

```{r}
knitr::kable(df)
```

# Task 3: List Practice

1.  Create df with placebo treatment:

```{r}
placebo_pre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

placebo_df <- data.frame(patient=paste("Subject", 1:10, sep = "_"),
                         pre_bp=placebo_pre,
                         post_bp=placebo_post,
                         diff_bp=placebo_pre-placebo_post
                         )

placebo_df$low_post_bp <- placebo_df$post_bp<120

placebo_df
```

2.  Create and store list containing each df:

```{r}
df_list <- list(treatment=df, placebo=placebo_df)

df_list
```

3.  Access the first list element using 3 types of syntax:

```{r}
df_list[1]

df_list[[1]]

df_list$treatment
```

4.  Using the list from (2), access *pre_bp* column of placebo df in one line of code:

```{r}
df_list$placebo[, 2, drop=FALSE]
```
